---
title: "Data Cleaning"
author: "Hannah Fazekas"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```

# Read in data

We are using data from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors =FALSE) 
head(catch_original)

```

pipes: %>%  cntl shift m 
df_cleaned <- df %>%  #df_cleaned gets df which gets the filter of a dataframe  
              filter(...) %>%  #which gets the select of that dataframe
              select(...)
# Clean and reshape
```{r}
#remove all column and notes column
catch_clean <- catch_original %>% 
      select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>%  # or select(-All, -notesRegCode)
      mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
      mutate(Chinook = as.numeric(Chinook))
     
head(catch_clean)
summary(catch_clean) #chinook is read as a character so we need to fix this (should be numbers)


```

## QA

Find where as.numeric could't convert Chinook to a number
```{r}
i <- which(is.na(catch_clean$Chinook))
i
catch_original[i, ]
```

Reshape from wide to tall format using 'pivot_longer'

```{r}
catch_long <- catch_clean %>% 
      pivot_longer(cols = -c(Region, Year), 
                      names_to = "species",
                      values_to = "catch") %>% 
          rename(catch_thousands = catch) %>% 
          mutate(catch = catch_thousands*1000) %>% 
          select(-catch_thousands)
         #mutate(catch= catch* 1000) = more elegant method to do this

head(catch_long)

```

```{r}
catch_wide <- catch_long %>% 
    pivot_wider(names_from = Year,
                  values_from = catch)
head(catch_wide)
```

## group_by and summarize

generate summay tables of your dataframes

```{r}
mean_region <- catch_long %>% 
      group_by(Region,species)  %>% # creates a semi-hidden attribute of your dataframe
      summarise(catch_mean = mean(catch), num_obs = n())

head(mean_region)

#in what year was the salmon catch the highest
annual_catch <- catch_long %>% 
      group_by(Year) %>% 
      summarise(catch_max = sum(catch)) %>% 
      arrange(-catch_max) #arrange dataframe in decreasing order

head(annual_catch)



```

What year had the highest sockeye catch

```{r}
annual_sockeye <- catch_long %>% 
      filter(species == "Sockeye") %>% 
      group_by(Year) %>% 
      summarise(catch_max = sum(catch)) %>% 
      arrange(-catch_max)
head(annual_sockeye)

```



# Join two data tables


Read in region data table

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea) #%>% 
    #rename(Region=code)

head(region_defs)
```


```{r}
catch_joined <- left_join(catch_long, region_defs, by= c("Region" = "code"))

head(catch_joined)

```

