---
title: "Geospatial Analysis"
author: "Hannah Fazekas"
date: "2/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(scales)
library(ggmap)
library(leaflet)
```

# Read in the data

```{r}
ak_regions <- read_sf("data/ak_regions_simp.shp")

plot(ak_regions) #does not look good

class(ak_regions) #sf object, and a data.frame, 

head(ak_regions) #geometry column -- geospatial vector data
#4326 is WGS84 coordinate systems (what is used by GPS)

st_crs(ak_regions) #function to print out what coordinate reference system you are using

```

coordinate reference systems are what you are using when making spatial maps
 
 map is bad because it is not a good coordinate reference system in terms of being recognized by r

contains the data and how to project that 3d data on a 2d surface

can look up coordinate systems on epsg.io 
We are using:
EPSG Codes:
3338 - Alaska Albers
4326 - WGS84 (GPS) 
3847 - pseudo meccator (Google Maps, Open Street Map)

# Convert coordinate systems 
  
  To more appropriately display map
  
```{r}
ak_regions_3338 <- ak_regions %>% 
      st_transform(crs = 3338)

st_crs(ak_regions_3338)

plot(ak_regions_3338)
```

# sf works with the tidyverse 

So, we can use all of those functions

```{r}
ak_regions_3338 %>% 
    filter(region == "Southeast") %>% 
    select(region)
    
```

The geometry column is 'sticky' and unless you explicitly tell R to drop it, it will continue to keep the information together. 


# Spatial joins and summaries

```{r}
pop <- read.csv("data/alaska_population.csv", stringsAsFactors = F)

head(pop)

#make pop a spatial object
pop_4326 <- st_as_sf(pop,
                     coords = c("lng", "lat"),
                     crs = 4326,
                     remove = FALSE) #we don't know but if it is lat long data, its a good bet that it is WGS84
head(pop_4326)

#use a spatial join to add pop data to region data
pop_joined <- st_join(pop_4326, ak_regions_3338, join=st_within)
#error because coordinate systems are not the same
pop_3338 <- pop_4326 %>% 
        st_transform(crs = 3338)

pop_joined <- st_join(pop_3338, ak_regions_3338, join=st_within)

head(pop_joined)

plot(pop_joined)
```

In st_join, there is an argument 'join' that requires you to specify a predicate function depending on your data. We have population points to match with polygons of region.
x and y objects must also be of class sf
We want to use the st_within

# Calculate population by region
```{r}
pop_region <- pop_joined %>% 
    group_by(region) %>% 
    summarise(total_pop = sum(population))

plot(pop_region) #keeps geometry so it doesn't really do what we want

?sf::tidyverse

pop_region <- pop_joined %>% 
    as.data.frame() %>% 
    group_by(region) %>% 
    summarise(total_pop = sum(population))

head(pop_region)
#rejoin this to shapefile

pop_region_3338 <- left_join(ak_regions_3338, pop_region)

plot(pop_region_3338)

```


#Example using tidyverse on spatial data with purpose
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
    group_by(mgmt_area) %>% 
    summarise(total_pop = sum(total_pop))
plot(pop_mgmt_3338["total_pop"])

```

```{r}
write_sf(pop_region_3338, "data/ak_regions_pop.shp", delete_layer = TRUE)
#delete_layer -- can remove layers 

```



# Make maps!

sf works really well with ggplot2

```{r}
rivers_3338 <- read_sf("data/ak_rivers_simp.shp")
st_crs(rivers_3338)

ggplot() + 
  geom_sf(data= pop_region_3338, aes(fill = total_pop))+ theme_bw()+
  geom_sf(data = pop_3338, aes(), size=0.5)+
  geom_sf(data=rivers_3338, aes(size= StrOrder), color="blue")+
  scale_size(range=c(0.01,0.2), guide = F)+
  labs(fill= "Total Population")+
  scale_fill_continuous(low = "khaki", high="firebrick", labels = comma)
  


```


# Getting basemaps using ggmap

Grabs stuff from Google Maps and Open Street Map

```{r}
pop_3857 <- pop_3338 %>% 
    st_transform(crs = 3857)

```

Going to get map from a server called stamenmap

```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

```{r}
bbox <- c(-170, 52, -130, 64) #lat longs in units of degrees
ak_map <- get_stamenmap(bbox, zoom = 4)
ak_map_3857 <- ggmap_bbox_to_3857(ak_map)

class(ak_map_3857) #special ggmap raster
plot(ak_map_3857)
```

```{r}
ggmap(ak_map_3857)+
  geom_sf(data = pop_3857, aes(color= population), inherit.aes = F)+
  scale_color_continuous(low="khaki", high = "firebrick", labels = comma)

```


## Interactive Maps using leaflet

```{r}
#define the leaflet projection
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))

st_crs(pop_region_3338)

pop_region_4326 <- pop_region_3338 %>% st_transform(crs = 4326)

```

```{r}

#leaflet uses pipes, not pluses like in ggplot
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = "gray",
                    weight = 1)

m

pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")


m
```


```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        #addTiles() %>% # to add world tiles. Must remove above arguments for it to match            the projections
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1) %>% 
        addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```

